2H qSIP Calculations

### 2021-10-27 qSIP_calculations.Rmd
fixed issue where at2H was parsed as chr

@@

# no zero samples
peaks_mapped_with_zeros %>% filter(is.na(zero_sample_id)) %>% View()

@@

```{r}
gen_calc_data <- peaks_mapped_with_zeros_analytes %>% 
  mutate(
    f_label = 0.5, # label is 0.5 at%
    a = 0.5, # assimilation efficiency
    t0 = 0, # start time
    f_start = at2H_mn_zero,
    d_t = inc_time_d - t0, # change in time
    f_t = at2H, # restate 2F at time t
    u_d = gencalc(a = 0.5, # Assimilation efficiency
                  FT = f_t, # 2F at time t (at%)
                  F0 = f_start, # 2F at t0 (at%)
                  FL = f_label, # 2F of label (at%)
                  t = d_t), # change in time
    # "negative" turnovers are artifacts of noise and should be zero
    u_d = if_else(u_d < 0, 0, u_d),
    # calculate generation time, converting negatives to zero
    gen_d = if_else(u_d < 0, 0, log(2) / u_d)
  ) %>%
  mutate(
    chain_fctr = compound %>% 
      str_remove("\\d+-OH") %>% 
      parse_number() %>% 
      abs() %>% 
      {paste0("C", .)} %>% 
      factor()
      ) %>% 
  mutate(
    chain_num = compound %>% 
      str_remove("\\d+-OH") %>% 
      parse_number() %>% 
      abs()
  )

@@

gen_calc_data_abbr <- gen_calc_data %>% 
  select(Analysis, sample_id, soil_id, terrain, replicate_id, compound, peak_nr, area2, 
         at2H, at2H_mn_zero,u_d, gen_d)
```

@@

```{r}
gen_calc_data %>% filter(inc_time_d != 0) %>% 
  mutate(compound = fct_reorder(compound, rt)) %>% 
  ggplot(aes(x = compound,
             y = u_d, 
             color = terrain, 
             fill = terrain)) +
  facet_wrap(vars(terrain, inc_time_d_str),
             ncol = 2) +
  geom_segment(aes(xend = compound,
                   yend = 0),
               color = "black") +
  geom_point() +
  scale_color_npg() +
  labs(x = "Compound",
       y = "Turnover time (d^-1)",
       title = "Compound-specific turnover rates of bacterial phospholipid fatty acids",
       color = "Terrain",
       fill = "Terrain") +
  theme_bw() +
  theme(
    axis.text.x = element_text(angle = 90, 
                               hjust = 1)
  )
```

@@

gen_calc_data %>% filter(inc_time_d != 0) %>% 
  ggplot(aes(x = rel_area_FID,
             y = u_d,
             color = chain_fctr)) +
  geom_point() +
  scale_color_viridis_d() +
  facet_wrap(~terrain) +
  labs(x = "Relative Area (GC-FID)",
       y = "Turnover (d^-1)",
       color = "Compound class") +
  theme_bw()

>>>>

# zero_peaks_averaged <- read_xlsx(file.path("data", "IRMS", "zero_peaks_averaged.xlsx"))

@@

# no zero samples
# peaks_mapped_with_zeros %>% filter(is.na(zero_sample_id)) %>% View()

@@

```{r}
gen_calc_data <- peaks_mapped_with_zeros_analytes %>% 
  mutate(
    f_label = 0.5, # label is 0.5 at%
    a = 0.5, # assimilation efficiency
    t0 = 0, # start time
    f_start = at2H_mn_zero,
    d_t = inc_time_d - t0, # change in time
    f_t = at2H, # restate 2F at time t
    u_d = gencalc(a = 0.5, # Assimilation efficiency
                  FT = f_t, # 2F at time t (at%)
                  F0 = f_start, # 2F at t0 (at%)
                  FL = f_label, # 2F of label (at%)
                  t = d_t), # change in time
    # "negative" turnovers are artifacts of noise and should be zero
    u_d = if_else(u_d < 0, 0, u_d),
    # calculate generation time, converting negatives to zero
    gen_d = if_else(u_d < 0, 0, log(2) / u_d)
  ) %>%
  mutate(
    chain_fctr = compound %>% 
      str_remove("\\d+-OH") %>% 
      parse_number() %>% 
      abs() %>% 
      {paste0("C", .)} %>% 
      factor()
      ) %>% 
  mutate(
    chain_num = compound %>% 
      str_remove("\\d+-OH") %>% 
      parse_number() %>% 
      abs()
  )
```

@@

# gen_calc_data_summarized %>% openxlsx::write.xlsx(file.path("data", "gen_calc_data_summarized.xlsx"))
```

@@

```{r}
gen_calc_data_summarized %>% 
  filter(inc_time_d != 0, compound != "i-16:0") %>%
  group_by(compound) %>% 
  summarize_if(is.numeric, mean, na.rm = TRUE) %>% 
  mutate(compound = fct_reorder(compound, u_d)) %>% 
  ggplot(aes(x = compound,
             y = u_d,
             fill = u_d)) +
  geom_col(position = "dodge") +
  scale_fill_viridis_c(option = "plasma") +
  labs(x = "Compound",
       y = "Mean Turnover time (d^-1)",
       title = "Compound-specific turnover rates of bacterial phospholipid fatty acids") +
  theme_bw() +
  theme(
    axis.text.x = element_text(angle = 90, 
                               hjust = 1)
  )
```

@@

gen_calc_data %>% 
  filter(inc_time_d != 0) %>% 
  ggplot(aes(x = rel_area_FID,
             y = u_d)) +
  geom_point(aes(color = chain_fctr)) +
  geom_smooth(method = "lm",
              se = FALSE) +
  scale_color_viridis_d() +
  facet_wrap(~terrain) +
  labs(x = "Relative Area (GC-FID)",
       y = "Turnover (d^-1)",
       color = "Compound class") +
  theme_bw()
